# Project Organization


## Project Structure I: Separating Inputs and Outputs

Structuring our project and the locations of files is an important concept.

Let's look at the structure of our project's folder.
Open a terminal and change into this directory

```{r cd-bash, engine = 'bash', eval = FALSE}
cd YOUR/PATH/TO/snakemake-econ-r-student
```

And list the subdirectories of the main directory

```{r ls-bash, engine = 'bash', eval = FALSE}
ls -d */
```

We see the following folder structure
```{r ls-bash-res, engine = 'bash', eval = FALSE}
./
    |- src/
    |- out/
    |- log/
    |- sandbox/
```


We recommend the following structure for any project:

* Root Folder
* `src` folder for input files
* `out` folder for output files
* a `log` folder to store computer logs
* a `sandbox` folder that gives us a 'safe place' to develop new code

We discuss each of these in turn.

### The Root Folder

TBD

### The `src` folder

TBD

### The `out` folder

TBD

### The `log` folder

### Exploring the Full Structure of the MRW Replication Project

Now, let's look at all contents of this main projects directory:

```{r ls-bash-2, engine = 'bash', eval = FALSE}
ls -F .
```

We see the following folder structure
```{r ls-bash-res-2, engine = 'bash', eval = FALSE}
./
    |- src/
    |- out/
    |- log/
    |- sandbox/
    | README.md
    | Snakefile
```

Notice that there are no instances of:
    (i) scripts,
    (ii) files containing content of the paper or slides
    (iii) something else we haven't thought of yet
Instead, there are only two files, a `README.md` and a file called `Snakefile.`

TODO: explain these two files

## Project Structure II: Separating Logical Chunks of the Project

TBD - disciplined subdirectory structure

### Exploring the Structure of the MRW Replication Subdirectories

TBD - openly discuss the structure of our working example

## Project Structure III: Separating Input Parameters from Code

TBD

### Exploring Parameter Separation in the MRW Replication Project
