# Target Rules

## Where we are now?

If you still have the `hello_world` rule in your Snakefile, now is a good moment to remove it.
Then, your Snakefile should look something like this:

```{r, engine = 'bash', eval = FALSE}
## Snakemake - MRW Replication
##
## @yourname


# --- Build Rules --- #

rule solow_intermediate:
    input:
        script   = "src/analysis/estimate_ols_model.R",
        data     = "out/data/mrw_complete.csv",
        model    = "src/model-specs/model_solow.json",
        subset   = "src/data-specs/subset_intermediate.json"
    output:
        estimate = "out/analysis/model_solow_subset_intermediate.rds",
    shell:
        "Rscript {input.script} \
            --data {input.data} \
            --model {input.model} \
            --subset {input.subset} \
            --out {output.estimate}"

rule solow_nonoil:
    input:
        script   = "src/analysis/estimate_ols_model.R",
        data     = "out/data/mrw_complete.csv",
        model    = "src/model-specs/model_solow.json",
        subset   = "src/data-specs/subset_nonoil.json"
    output:
        estimate = "out/analysis/model_solow_subset_nonoil.rds",
    shell:
        "Rscript {input.script} \
            --data {input.data} \
            --model {input.model} \
            --subset {input.subset} \
            --out {output.estimate}"

rule solow_oecd:
    input:
        script   = "src/analysis/estimate_ols_model.R",
        data     = "out/data/mrw_complete.csv",
        model    = "src/model-specs/model_solow.json",
        subset   = "src/data-specs/subset_oecd.json"
    output:
        estimate = "out/analysis/model_solow_subset_oecd.rds",
    shell:
        "Rscript {input.script} \
            --data {input.data} \
            --model {input.model} \
            --subset {input.subset} \
            --out {output.estimate}"

rule gen_regression_vars:
    input:
        script = "src/data-management/gen_reg_vars.R",
        data   = "out/data/mrw_renamed.csv"
    output:
        data   = "out/data/mrw_complete.csv"
    shell:
        "Rscript {input.script} \
            --data {input.data} \
            --out {output.data}"

rule rename_vars:
    input:
        script = "src/data-management/rename_variables.R",
        data   = "src/data/mrw.dta"
    output:
        data = "out/data/mrw_renamed.csv"
    shell:
        "Rscript {input.script} \
            --data {input.data} \
            --out {output.data}"

# --- Clean Rules --- #

rule clean:
    shell:
        "rm -rf out/*"

rule clean_data:
    shell:
        "rm -rf out/data/*"

rule clean_analysis:
    shell:
        "rm -rf out/analysis/*"

```

## How Snakemake determines the build order when multiple rules are present

Now that we have worked with multiple rules and seen how rules can call other rules, let us try to understand the principle behind this.

As we know, we can execute any rule in `Snakefile` explicitly by calling it by name like so:

```{r, engine = 'bash', eval = FALSE}
snakemake --cores 1 solow_intermediate
```

When no rule name is explicitly given, Snakemake will execute the first rule it encounters at the top of `Snakefile`:

```{r, engine = 'bash', eval = FALSE}
snakemake --cores 1
```

If you followed the order of rules in the solution to the previous exercise, `solow_intermediate` is the first rule in `Snakefile`.
In this case, command are equivalent.

The rule we ask `Snakemkake` to execute either explicitly or implicitly is called a **target rule**.

If all necessary inputs to build the target rule exist, Snakemake will execute the rule and build the defined outputs.

When Snakemake recognizes that a necessary `input` to execute the target rule is missing, Snakemake will try to build it through the other rules present in `Snakefile`.

To illustrate this, let us stick to our present project and assume the target rule is `solow_intermediate`.
The following graph shows all rules which lead up to the target rule:

![filegraph](img/filegraph.pdf)

`mrw_complete.csv` is a necessary input to `solow_intermediate`.
If the file does not exist, `Snakemake` will search through `Snakefile` for another rule which has `mrw_complete.csv` as its output -- in our case `gen_regression_vars`.
In this case, Snakemake will first execute `gen_regression_vars` before executing the target rule `solow_intermediate`.

If necessary inputs are missing to execute `gen_regression_vars`, Snakemake will search for other rules to produce it and so forth...

To allow Snakemake to do work properly, the input and output relationships in a project need to follow a **directed acyclic graph (DAG)**.
Directed here means that....
Acyclic implies that there can be no two rules which require output from one another as inputs, either directly or in a larger loop.

This has a few implications on how we should define our intermediate input and output files.
We should follow the following rules to prevent problems:

- **No rule should have use the same file as input and output.** It is often common practice to load a certain dataset, perform operations on it and overwrite the original file. While there are good reasons to never do this in any workflow, this behavior can lead to more severe problems when using Snakemake. Snakemake only seaches for rules which can create input files which do not yet exist. If a rule overwrites an already existing input file, Snakemake would not execute it as a dependency for a downstream target rule and the changes of such a rule would not be reflected in the outputs of the target rule. For this reason, `gen_regression_vars` writes its results to a new file in this project.
- **No two rules should have the same output.** When Snakemake searches for a file to create a necessary input, it will execute the first rule it encounters which produces the file. To make input and output relationships explicit and reproducible, we do not want to let this be determined by chance.
- **A project should have a clear direction.** The projects needs to be directed which implies that the execution of a project should have a clear trajetory. In practice this typically starts with data manipulations which are necessary to perform analysis, which in turn will be used to create plots and tables, which finally end up in a paper or set of slides.
Directedness here for example implies that the regression table cannot be an input to data manipulations etc.
- **A project cannot be circular.** Any circularity would let Snakemake search for rules in an infinite loop.

## Dedicated target rules to execute multiple rules

