# Make Tables

TODO: Add current state of Snakefile

Now that we have learned how to make our code more concise, we want to move forward and apply the new concepts from the previous part to finish our replication.

In this chapter we will create a rule which generates all the regression tables in @mrw.

Let us first start by making a single, explicit rule to create our first table.

You can find the R script which does this under `src/tables/regression_table.R`.
As always, we start by using the help function to learn the necessary input parameters the script needs:

```{r, engine = 'bash', eval = FALSE}
Rscript src/table/regression_table.R --help
```

```{r, engine = 'bash', eval = FALSE}
Usage: src/tables/regression_table.R [options]

Options:
        -s CHARACTER, --spec=CHARACTER
                a json dictionary of table parameters

        -o CHARACTER, --out=CHARACTER
                output file name [default = out.tex]

        -h, --help
                Show this help message and exit
```

Our model expects two arguments:

- `--spec`, a json dictionary with table parameters. This includes the models this tables contains as well as table specifications such as headers, column names etc.
- `--out`, the output filename the \LaTeX code will be saved to

We will make the first table with `table_01.json` as the `spec` argument:

```{r, engine = 'bash', eval = FALSE}
rule table:
    input:
        script = "src/tables/regression_table.R",
        spec   = "src/table-specs/table_01.json"
    output:
        table = "out/tables/table_01.tex"
    shell:
        "Rscript {input.script} \
            --spec {input.spec} \
            --out {output.table}"           
```

If all models files have been created in the last chapter, then this rule should run through with a single executed task.
Additional to the usual Snakemake output, the \LaTeX code of the table is printed to the screen.

Let us check that the table has been successfully created in the `out/tables` folder:

```{r, engine = 'bash', eval = FALSE}
ls out/tables
```

```{r, engine = 'bash', eval = FALSE}
table_01.tex
```

Nice, let us now clean our output folder and try to execute the table rule again.

```{r, engine = 'bash', eval = FALSE}
snakemake --cores 1 clean
snakemake --cores 1 table
```

TODO: Recreate on English system

```{r, engine = 'bash', eval = FALSE}
Building DAG of jobs...
Using shell: /bin/bash
Provided cores: 4
Rules claiming more threads will be scaled down.
Job counts:
	count	jobs
	1	table
	1

[Sat Feb 20 21:19:20 2021]
rule table:
    input: src/tables/regression_table.R, src/table-specs/table_01.json
    output: out/tables/table_01.tex
    jobid: 0

[...]

In gzfile(file, "rb") :
  kann komprimierte Datei '/Users/ubergm/Documents/projects/snakemake/snakemake-econ-r-learner/out/analysis/model_solow.subset_nonoil.rds' nicht öffnen. Grund evtl. 'No such file or directory'
Ausführung angehalten

[Sat Feb 20 21:19:20 2021]
Error in rule table:
    jobid: 0
    output: out/tables/table_01.tex
    shell:
        Rscript src/tables/regression_table.R             --spec src/table-specs/table_01.json             --out out/tables/table_01.tex
        (one of the commands exited with non-zero exit code; note that snakemake uses bash strict mode!)

Shutting down, this might take some time.
Exiting because a job execution failed. Look above for error message
Complete log: /Users/ubergm/Documents/projects/snakemake/snakemake-econ-r-learner/.snakemake/log/2021-02-20T211920.004018.snakemake.log
```

We got an error.
Looking at the R output, we see that the file `out/analysis/model_solow.subset_nonoil.rds` could not be found.
Why is this?

The model specification includes the path to the necessary models featured in table 1.
We have not included these models as necessary inputs.

This is easy to fix.
We just need to include the model inputs from the `run_models` target rule from the previous chapter into our new rule.
This tells Snakemake to create all models before creating the regression table.

Copying this part will give us a rule like this:

```{r, engine = 'bash', eval = FALSE}
rule table:
    input:
        script = "src/tables/regression_table.R",
        spec   = "src/table-specs/table_01.json",
        models = expand("out/analysis/{iModel}.{iSubset}.rds",
                        iModel = MODELS,
                        iSubset = DATA_SUBSET)        
    output:
        table = "out/tables/table_01.tex"
    shell:
        "Rscript {input.script} \
            --spec {input.spec} \
            --out {output.table}"           
```

Let us clean and rerun the rule:

```{r, engine = 'bash', eval = FALSE}
snakemake --cores 1 clean
snakemake --cores 1 table
```

Checking the output folder again, we see that all models

```{r, engine = 'bash', eval = FALSE}
ls out/analysis
```

```{r, engine = 'bash', eval = FALSE}
model_aug_cc.subset_intermediate.rds	    model_aug_cc_restr.subset_nonoil.rds     model_aug_solow.subset_oecd.rds		    model_cc.subset_intermediate.rds	 model_solow.subset_nonoil.rds		    model_solow_restr.subset_oecd.rds
model_aug_cc.subset_nonoil.rds		    model_aug_cc_restr.subset_oecd.rds	     model_aug_solow_restr.subset_intermediate.rds  model_cc.subset_nonoil.rds		 model_solow.subset_oecd.rds		    model_ucc.subset_intermediate.rds
model_aug_cc.subset_oecd.rds		    model_aug_solow.subset_intermediate.rds  model_aug_solow_restr.subset_nonoil.rds	    model_cc.subset_oecd.rds		 model_solow_restr.subset_intermediate.rds  model_ucc.subset_nonoil.rds
model_aug_cc_restr.subset_intermediate.rds  model_aug_solow.subset_nonoil.rds	     model_aug_solow_restr.subset_oecd.rds	    model_solow.subset_intermediate.rds  model_solow_restr.subset_nonoil.rds	    model_ucc.subset_oecd.rds
```

and the table

```{r, engine = 'bash', eval = FALSE}
ls out/tables
```

```{r, engine = 'bash', eval = FALSE}
table_01.tex
```

have been successfully created.

We want to generalize this in the following exercise to automatically create all tables from @mrw.

### Exercise {- .exercise}

1. Add a wildcard to the `table` rule which can take the varying part of the `{input.spec}` and `{output.table}` parts. To do so, replace `table_01` with `{iTable}` in both.
2. Create a new target rule called `make_tables`. The rule expands the output argument of the `table` rule. It uses the `{iTable}` wildcard from a list called `TABLES`.
3. Create the TABLES list using the `glob_wildcards function` which matches filename from the `src/table-spec` folder.
4. Clean your output folder.
5. Run the `make_tables` rule.
6. Verify that all six tables have been created in the output folder.

### Solution {- .solution}

Putting Steps 1-3 together should give us these three new parts:

```{r, engine = 'bash', eval = FALSE}
TABLES = glob_wildcards("src/table-specs/{fname}.json").fname

rule make_tables:
    input:
        expand("out/tables/{iTable}.tex",
                iTable = TABLES)

rule table:
    input:
        script = "src/tables/regression_table.R",
        spec   = "src/table-specs/{iTable}.json",
        models = expand("out/analysis/{iModel}.{iSubset}.rds",
                        iModel = MODELS,
                        iSubset = DATA_SUBSET),
    output:
        table = "out/tables/{iTable}.tex"
    shell:
        "Rscript {input.script} \
            --spec {input.spec} \
            --out {output.table}"

[...]
```

4. Clean your output folder.

```{r, engine = 'bash', eval = FALSE}
snakemake --cores 1 clean
```

5. Run the `make_tables` rule.

```{r, engine = 'bash', eval = FALSE}
snakemake --cores 1 make_tables
```

6. Verify that all six tables have been created in the output folder.

```{r, engine = 'bash', eval = FALSE}
ls out/tables
```

The expected output are the six table \LaTeX files:

```{r, engine = 'bash', eval = FALSE}
table_01.tex  table_02.tex  table_03.tex  table_04.tex	table_05.tex  table_06.tex
```