# Config Files

Motivation...

## The `config.yaml` filepath

In our project's root directory we have the file `config.yaml`:

```{r, engine = 'bash', eval = FALSE}
$ ls -F
```
```{r, engine = 'out', eval = FALSE}
config.yaml  find_r_packages.sh  HELP.txt
install_r_packages.R  logs/  out/
README.md  REQUIREMENTS.txt  sandbox/
Snakefile  src/
```

If we look at the contents of `config.yaml` we see a collection of paths that
    match up to the paths in our project:
```{r, engine = 'bash', eval = FALSE}
$ cat config.yaml
```
```{r, engine = 'out', eval = FALSE}
ROOT: "."
sub2root: "../../"
src: "src/"
log: "logs/"
out: "out/"
src_data: "src/data/"
src_data_mgt: "src/data-management/"
src_analysis: "src/analysis/"
src_lib: "src/lib/"
src_model_specs: "src/model-specs/"
src_data_specs: "src/data-specs/"
src_tables: "src/tables/"
src_figures: "src/figures/"
src_paper: "src/paper/"
src_slides: "src/slides/"

out_analysis: "out/analysis/"
out_data: "out/data/"
out_figures: "out/figures/"
out_tables: "out/tables/"
out_paper: "out/paper/"
out_slides: "out/slides/"
```

We can use this collection of paths to simplify the content of the Snakefile.
First we must import the `config.yaml` file into the Snakefile.
We import it using the `configfile:` notation, and so it's easy for us to find,
    we place it at the very top of our Snakefile.

```{r, engine = 'bash', eval = FALSE}
## Snakemake - MRW Replication
##
## @yourname
##

# --- Importing Configuration Files --- #

configfile: "config.yaml"

# --- Dictionaries --- #
<...>
```

## Using paths from the config file

We can now use the paths in `config.yaml` throughout the Snakefile.
The text to the left of the colon in each line of the file is our reference to a
    particular path, so `src_data` will serve as our reference to the folder `src/data/`.
Snakemake does not knows this path reference comes from the config file unless we instruct it to,
    so we reference it as `config["src_data"]`.
The path is then connected to the filename with a `+`.
Hence we can reference the original MRW data located in `src/data` as `config["src_data"] + "mrw.dta"`
Here is the new version of the rule `rename_vars` using the config file to simplify the paths:

```{r, engine = 'bash', eval = FALSE}
## rename_vars        : creates meaningful variable names
rule rename_vars:
    input:
        script = config["src_data_mgt"] + "rename_variables.R",
        data   = config["src_data"] + "mrw.dta"
    output:
        data = config["out_data"] + "mrw_renamed.csv"
    log:
        config["out_log"] + "data_mgt/rename_variables.Rout"
    shell:
        "Rscript {input.script} \
            --data {input.data} \
            --out {output.data} \
            >& {log}"
```
