# Adding Parameters

In our analysis pipeline so far we have always passed files to our R scripts.
While we promote this practice, and wrapping important information into .json files,
    sometimes it seems like overkill to write a new json file to contain one line of configuration.
An alternative to passing these json files is to use Snakemake's built in
    `params` arguments, which are rule specific,
    to store information that we want to pass to our R Script.
This goal of this chapter is to show how to use `params` to pass across a piece of information.

## Motivating Example: Constructing a Regression Table from OLS Results

So far, we have estimated a series of OLS regressions and stored there output inside the `out/analysis` directory.
Typically once we have estimated one or more models, we want to format the output into a regression table that we can insert into a written document like a paper or set of presentation slides.
In the folder `src/tables/` we can see that there are a series of R scripts:
```{r, engine = 'bash', eval = FALSE}
$ ls src/tables/
```
which prints to the screen:
```{r, engine = 'out', eval = FALSE}
tab01_textbook_solow.R  tab03_ucc_solow.R  tab05_cc_aug_solow.R
tab02_augment_solow.R   tab04_cc_solow.R   tab06_cc_aug_solow_restr.R
```
This shows that the example is designed to build 6 tables.
Each table has it's own script that constructs it.
We will start by constructing Table 1, from `tab01_textbook_solow.R`
Let's have a look at what information this script expects us to pass using the help flag:

```{r, engine = 'bash', eval = FALSE}
Rscript src/tables/tab01_textbook_solow.R --help
```
```{r, engine = 'out', eval = FALSE}
Usage: src/tables/tab01_textbook_solow.R [options]


Options:
	-fp CHARACTER, --filepath=CHARACTER
		A directory path where models are saved

	-m CHARACTER, --models=CHARACTER
		A regex of the models to load

	-o CHARACTER, --out=CHARACTER
		output file name [default = out.tex]

	-h, --help
		Show this help message and exit
```

From this we learn that we need to pass:

1. `--filepath`, which is the directory where our OLS models are stored
2. `--models`, a regular expression to tell R which models within the filepath to workwith
3. `--out`, a .tex file where we want to direct the output

Now we will work on constructing this rule.

## Creating a Rule with `params`

We are going to use the `params` option to pass across the filepath and the models regular expression into R.
A sketch of the rule we want to create is:

```{r, engine = 'bash', eval = FALSE}
rule textbook_solow:
    input:
        script = ,
        models =
    params:
        filepath   = ,
        model_expr =
    output:
        table = ,
    shell:
        "Rscript {input.script} \
            --filepath {params.filepath} \
            --models {params.model_expr} \
            --out {output.table}"
```

There are two important points to notice about how we added params to our rules:

1. `params` are added to the rule in a similar way to inputs and lists
2. `params` are referenced identically to inputs and outputs in the shell command

Now we need to decide what information needs to be entered into each line of our rule.
EXPLAIN

Our rule then becomes:

```{r, engine = 'bash', eval = FALSE}
rule textbook_solow:
    input:
        script = "src/tables/tab01_textbook_solow.R",
        models = expand("out/analysis/{iModel}_ols_{iSubset}.rds",
                            iModel = MODELS,
                            iSubset = DATA_SUBSET)),
    params:
        filepath   = "out/analysis/",
        model_expr = "model_solow*.rds"
    output:
        table = "out/tables/tab01_textbook_solow.tex"
    shell:
        "Rscript {input.script} \
            --filepath {params.filepath} \
            --models {params.model_expr} \
            --out {output.table}"
```

There are two ways to run this rule:

1. Tell snakemake to run this rule explicitly, `snakemake textbook_solow`
    * Because it is not the first rule of the Snakefile it isnt run by default
2. Add the output of this rule to the `all` rule.
    * Adds creating this table to our complete analysis pipeline

We prefer (2).
Hence we also update the `all` rule as follows:

```{r, engine = 'bash', eval = FALSE}
rule all:
    input:
        figs   = expand("out/figures/{iFigure}.pdf",
                            iFigure = FIGURES),
        models = expand("out/analysis/{iModel}_ols_{iSubset}.rds",
                            iModel = MODELS,
                            iSubset = DATA_SUBSET),
        tab01  = "out/tables/tab01_textbook_solow.tex"
```

If we then do a dry run to see what Snakemake plans to do^[
    An alternative would be to run `snakemake --summary` and examine the output.
]:

```{r, engine = 'out', eval = FALSE}
Building DAG of jobs...
Job counts:
	count	jobs
	1	all
	1	textbook_solow
	2

[Tue Feb  5 17:42:59 2019]
rule textbook_solow:
    input: src/tables/tab01_textbook_solow.R, out/analysis/model_solow_ols_subset_oecd.rds, out/analysis/model_solow_ols_subset_nonoil.rds, out/analysis/model_solow_ols_subset_intermediate.rds, out/analysis/model_aug_cc_restr_ols_subset_oecd.rds, out/analysis/model_aug_cc_restr_ols_subset_nonoil.rds, out/analysis/model_aug_cc_restr_ols_subset_intermediate.rds, out/analysis/model_solow_restr_ols_subset_oecd.rds, out/analysis/model_solow_restr_ols_subset_nonoil.rds, out/analysis/model_solow_restr_ols_subset_intermediate.rds, out/analysis/model_cc_ols_subset_oecd.rds, out/analysis/model_cc_ols_subset_nonoil.rds, out/analysis/model_cc_ols_subset_intermediate.rds, out/analysis/model_ucc_ols_subset_oecd.rds, out/analysis/model_ucc_ols_subset_nonoil.rds, out/analysis/model_ucc_ols_subset_intermediate.rds, out/analysis/model_aug_solow_restr_ols_subset_oecd.rds, out/analysis/model_aug_solow_restr_ols_subset_nonoil.rds, out/analysis/model_aug_solow_restr_ols_subset_intermediate.rds, out/analysis/model_aug_cc_ols_subset_oecd.rds, out/analysis/model_aug_cc_ols_subset_nonoil.rds, out/analysis/model_aug_cc_ols_subset_intermediate.rds, out/analysis/model_aug_solow_ols_subset_oecd.rds, out/analysis/model_aug_solow_ols_subset_nonoil.rds, out/analysis/model_aug_solow_ols_subset_intermediate.rds
    output: out/tables/tab01_textbook_solow.tex
    jobid: 27


[Tue Feb  5 17:42:59 2019]
localrule all:
    input: out/figures/conditional_convergence.pdf, out/figures/unconditional_convergence.pdf, out/figures/aug_conditional_convergence.pdf, out/analysis/model_solow_ols_subset_oecd.rds, out/analysis/model_solow_ols_subset_nonoil.rds, out/analysis/model_solow_ols_subset_intermediate.rds, out/analysis/model_aug_cc_restr_ols_subset_oecd.rds, out/analysis/model_aug_cc_restr_ols_subset_nonoil.rds, out/analysis/model_aug_cc_restr_ols_subset_intermediate.rds, out/analysis/model_solow_restr_ols_subset_oecd.rds, out/analysis/model_solow_restr_ols_subset_nonoil.rds, out/analysis/model_solow_restr_ols_subset_intermediate.rds, out/analysis/model_cc_ols_subset_oecd.rds, out/analysis/model_cc_ols_subset_nonoil.rds, out/analysis/model_cc_ols_subset_intermediate.rds, out/analysis/model_ucc_ols_subset_oecd.rds, out/analysis/model_ucc_ols_subset_nonoil.rds, out/analysis/model_ucc_ols_subset_intermediate.rds, out/analysis/model_aug_solow_restr_ols_subset_oecd.rds, out/analysis/model_aug_solow_restr_ols_subset_nonoil.rds, out/analysis/model_aug_solow_restr_ols_subset_intermediate.rds, out/analysis/model_aug_cc_ols_subset_oecd.rds, out/analysis/model_aug_cc_ols_subset_nonoil.rds, out/analysis/model_aug_cc_ols_subset_intermediate.rds, out/analysis/model_aug_solow_ols_subset_oecd.rds, out/analysis/model_aug_solow_ols_subset_nonoil.rds, out/analysis/model_aug_solow_ols_subset_intermediate.rds, out/tables/tab01_textbook_solow.tex
    jobid: 0

Job counts:
	count	jobs
	1	all
	1	textbook_solow
	2
```
We see that snakemake only needs to create the table from our newly created rule.
Now run snakemake to build the table:
```{r, engine = 'bash', eval = FALSE}
$ snakemake all
```
and when finished if we list the contents of `out/tables` we see our new regression table has been created:
```{r, engine = 'bash', eval = FALSE}
$ ls out/tables/
```
```{r, engine = 'out', eval = FALSE}
tab01_textbook_solow.tex
```

### Exercise: Building Table 2 {- .exercise}

Using the same rule format as above, incorporate `params` into a new rule called `augment_solow` that constructs Table 2.
